Automatic accent annotation for Russian

### 2014-12-09 rusaccent-nodalida-2015/rusaccent.tex
the next draft will suck less
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@57952 72bbbca6-d526-0410-a7d9-f06f51895060

@@

\author{Robert Reynolds \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@uit.no} \\\And
  Francis M. Tyers \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@uit.no} \\}

@@

This document has been adapted from the instructions for earlier
(E)ACL-proceedings, including those for EACL-2014 by Gosse Bouma
and Yannick Parmentier, those for ACL-2013 by Roberto Navigli
and Jing-Shing Chang, those for ACL-2012 by Maggie Li and Michael
White, those for ACL-2010 by Jing-Shing Chang and Philipp Koehn,
those for ACL-2008 by Joakim Nivre and Noah Smith, 
those for EACL-2006 by Eneko Agirre and Sergi Balari, 
those for ACL-2005 by Hwee Tou Ng and Kemal Oflazer, those for
ACL-2002 by Eugene Charniak and Dekang Lin, and earlier ACL and EACL
formats. Those versions were written by several people, including John
Chen, Henry S. Thompson and Donald Walker. Additional elements were
taken from the formatting instructions of the {\em International Joint
  Conference on Artificial Intelligence}.

\section{Introduction}
% das: removed reference to PostScript

>>>>

%\usepackage[T2A,T1]{fontenc}
%\usepackage[russian,english]{babel}
%\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}

@@

\author{Author1 \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain.com} \\\And
  Author2 \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain.com} \\}

@@

The template had a credits section for the history of the template, but we don't need
a Credits section, right?

\section{Introduction}

Russian word stress is variable, which means that the position of the stress in a word is not 
generally predictable. Furthermore, the inflecting word classes include complex patterns of
shifting stress. Lexical stress and its attendant vowel reduction are a prominent feature of
spoken Russian, but standard written Russian does not typically mark word stress. Students
of Russian frequently struggle to know how to pronounce written words, so many
textbooks mark stress. Even texts intended for native speakers
sometimes mark words that cannot be disambiguated through context.

Word stress ambiguity in Russian arises both between lexemes and within inflectional
paradigms. Stress ambiguity between lexemes includes pairs of the type
????? \emph{zámok} `castle' and ????? \emph{zamók} 'lock'. For every inflected form of 
these two words, they are both morphosyntactically and orthographically identical, their
meaning differentiated only by stress placement. This kind of ambiguity is relatively uncommon
in Russian. This paper focuses on stress ambiguities which are connected to morphosyntactic
differences. ADD EXAMPLE HERE with ????????, ????????, ????????, AND ??????/??????.
ADD ANOTHER EXAMPLE WITH the paradigms of ???? and ??????.

\subsection{Related Work}

Automatic stress placement in Russian is similar to Diacritic Restoration, a task
which has received increasing interest over the last 20 years. Generally speaking, diacritics
disambiguate otherwise homographic wordforms, so missing diacritics can
complicate many NLP tasks, such as text-to-speech and speech-to-text. For example, speakers 
of Czech and French sometimes type emails and other communications
without standard diacritics. In order to generate speech from these texts, they must first be 
normalized by restoring diacritics. A slightly different
situation arises with Arabic and Hebrew, since standard orthography lacks vowel diacritics
altogether. For such languages, the `restoration' of vowel diacritics results in less
ambiguity than in standard orthography. For languages with inherently ambiguous orthography,
it may be more precise to refer to this task as `diacritic enhancement', since it produces
text that is less ambiguous than the standard language. In this sense, Russian orthography
is similar to Arabic and Hebrew.
